### 2023-02-15 script/rembrandt/rembrandt_pathways.Rmd
update scri^pt mega bug globalboosttest

---
title: "Rembrandt pathway analysis"
author: "TF"
date: "08/06/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r}
library(dplyr)
library(ggplot2)
library(survival)
```

```{r data, results='hide'}
dfRemb <- readRDS(file = "../../data/rembrandt/dfGenesAndClinical.rds")
pathways <- GSA::GSA.read.gmt("../../data/rembrandt/c6.all.v7.1.symbols.gmt")
```

```{r}
lapply(pathways$genesets,
       function(vec_genes){
         dfRemb %>% select(any_of(vec_genes)) %>% ncol()
       }) %>%
  unlist() %>%
  hist(main = "Number of genes by pathway found in Rembrandt data")

lapply(pathways$genesets,
       length) %>%
  unlist() %>%
  hist(main = "Number of genes by pathway")

```