### 2023-06-07 reporting/discordance LV TF.Rmd
update results

---
title: "Discordance LV - TF"
author: "TF"
date: "`r Sys.Date()`"
output:
  bookdown::github_document2:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# Plots

## Draft plot

```{r eval=TRUE, fig.cap="Original plot", fig.height=20, include = TRUE, out.height=700, out.width=700, echo=FALSE}
knitr::include_graphics("images/comparison_methods6-1.png")
```

# New plot (fails to replicate)

```{r eval=TRUE, fig.cap="New plot (fails to replicate)", fig.height=20, include = TRUE, out.height=700, out.width=700, echo=FALSE}
knitr::include_graphics("images/plot_original.png")
```

# Data generation

## Variance covariance

### TF

```{r}
if(case == 1){

  mat_var_covar <- matrix(data = 0, nrow = nb_genes, ncol = nb_genes)

} else if(case == 2){
  # varcovar for non significant genes
  mat_var_covar <- matrix(data = 0, nrow = nb_genes, ncol = nb_genes)

  # covar for significant genes
  nb_sig_gene = prop_sig_gene * nb_genes

  covar <- 0.4*rnorm(n = nb_sig_gene^2/2-nb_sig_gene/2,
                     mean = 0.4,
                     sd = 0.1)

  # fill upper triangle
  mat_var_covar[1:nb_sig_gene,1:nb_sig_gene][upper.tri(mat_var_covar[1:nb_sig_gene,1:nb_sig_gene])] <- covar
  # duplicate to lower triangle
  mat_var_covar[lower.tri(mat_var_covar)] <- t(mat_var_covar)[lower.tri(mat_var_covar)]

} else if(case == 3){
  # varcovar for significant genes
  covar <- 0.4*rnorm(n = nb_genes^2,
                     mean = 0,
                     sd = 0.01)
  mat_var_covar <- matrix(data = covar,
                          nrow = nb_genes,
                          ncol = nb_genes)
  # duplicate upper triangle to lower triangle
  mat_var_covar[lower.tri(mat_var_covar)] <- t(mat_var_covar)[lower.tri(mat_var_covar)]

} else {
  stop("case must be 1, 2 or 3")
}

# set the variance
diag(mat_var_covar) <- variance

return(mat_var_covar)
```

### LV

```{r}
xsigma<-diag(rep(0.2,ng)) # The variance corresponding to Cjj ?

if (corr==2) {
  xgs<-diag(rep(0.2,ngs))        
  xgs[lower.tri(xgs==TRUE)]<-0.4*(rnorm(ncov,0.4,0.1))
  tmp<-xgs+t(xgs)-diag(diag(xgs))
  xsigma[1:ngs,1:ngs]<-tmp
}
if (corr==3) {
  xgs<-diag(rep(0.2,ngs))        
  xgs[lower.tri(xgs==TRUE)]<-0.4*sample(c(-1,1), size = 1, prob = c(1/2,  1/2),replace = TRUE)*(rnorm(ncov,0.4,0.1))
  tmp<-xgs+t(xgs)-diag(diag(xgs))
  xsigma[1:ngs,1:ngs]<-tmp
}
if (corr==4) {
  xgs<-diag(rep(0.2,ng))        
  xgs[lower.tri(xgs==TRUE)]<-0.4*(rnorm(ncov2,0,0.01))
  tmp<-xgs+t(xgs)-diag(diag(xgs))
  xsigma[1:ng,1:ng]<-tmp
}       
x<-rmvnorm(n=ns,mean=xmean,sigma=xsigma,method="chol")
```

### Comparison

-   Diagonal variance is 0.2 in both
-   Covariance corr is not 2,3,4 in LV is similar to case == 1 in TF
-   corr == 2 in LV is similar to case == 2 in TF
-   corr == 4 in LV is similar to case == 3 in TF

## Beta

### TF

```{r}
nb_sig_gene = round(prop_sig_gene*nb_genes)

if(type == "A"){

  res <- c(rnorm(n = nb_sig_gene, mean = 0, sd = 0.5),
           rep(0, nb_genes-nb_sig_gene))

} else if(type == "B"){

  res <- c(rnorm(n = ceiling(nb_sig_gene/2), mean = 0.1, sd = 0.5),
           rnorm(n = floor(nb_sig_gene/2), mean = -0.1, sd = 0.5),
           rep(0, nb_genes-nb_sig_gene))

} else if(type == "C"){

  res <- c(rnorm(n = ceiling(nb_sig_gene/2), mean = 0.1, sd = 0.25),
           rnorm(n = floor(nb_sig_gene/2), mean = -0.1, sd = 1),
           rep(0, nb_genes-nb_sig_gene))

} else {
  stop("case must be A, B or C")
}

return(res)
```

### LV

```{r}
if (params[ite,1]==0.1){
  gsp=0.3
}
if (params[ite,1]==0.3){
  gsp=0.4
}

if (params[ite,1]==0.5){
  gsp=0.5
}
gsize<-50 #gene set size
#gsp<-agsp[sigp]
ngs<-gsize*gsp
tau <- 0.5

if (beta==1) {
  bmean<-c(rep(0.1,floor(ngs/2)),
           rep(-0.5,(ngs-floor(ngs/2))),
           rep(0,ng-ngs)) #for power
  bsd<-c(rep(tau,floor(ngs/2)),
         rep(tau/2,(ngs-floor(ngs/2))),
         rep(0,ng-ngs)) #for power
}else{
  bmean<-c(rep(0.1,floor(ngs/2)),
           rep(-0.1,(ngs-floor(ngs/2))),
           rep(0,ng-ngs)) #for power
  bsd<-c(rep(tau/2,floor(ngs/2)),
         rep(tau*2,(ngs-floor(ngs/2))),
         rep(0,ng-ngs)) #for power
}

tbeta<-rnorm(ng,mean=bmean,sd=bsd)
```

### Comparison

-   Nothing corresponds to case == 'A' in LV script
-   There is no scenario in the paper where beta follows gaussian distribution with mean -0.5 (beta == 1 in LV script)
-   The variance defined in LV when beta != 1 corresponds to type == "C" in TF (because tau = 0.5)
- The proportion of significant gene does not match the ones in the paper

## Survival

### TF

```{r}
predictor<-x%*%vec_beta
time <- simu_simple_beta(x = x, predictor = predictor, slam=slam)

if (censoring==0) {
  df_survival <- data.frame(time = time,
                            event = 1)
}else{
  clam<-exp(predictor)*slam*censoring/(1-censoring)
  csg<-rexp(length(vec_beta),rate=clam)
  dim(csg)<-c(length(vec_beta),1)
  csgind<-(time<=csg)
  otime<-time*csgind+csg*(1-csgind)

  df_survival <- data.frame(time = otime,
                            event = as.numeric(csgind))
}

simu_simple_beta <- function(predictor,
                             slam = 0.005,
                             x){
  nb_observations <- nrow(x)
  randu<-runif(nb_observations,min=0,max=1)
  dim(randu)<-c(nb_observations,1)
  time<-exp(-predictor)*(-log(1-randu))/slam
  return(time)
}

```

### LV

```{r}
tbeta<-rnorm(ng,mean=bmean,sd=bsd)
predictor<-x%*%tbeta
randu<-runif(ns,min=0,max=1)
dim(randu)<-c(ns,1)
slam=0.005
time<-exp(-predictor)*(-log(1-randu))/slam
if (cp==0) {
  otime<-time
  csgind<-rep(1,ns)
}else{
  clam<-exp(predictor)*slam*cp/(1-cp)
  csg<-rexp(ns,rate=clam)
  dim(csg)<-c(ns,1)
  csgind<-(time<=csg)
  otime<-time*csgind+csg*(1-csgind)
}
```

### Comparison

Same implementation