### 2023-06-21 reporting/simu_rembrandt_breast.Rmd
rename varcovar

---
title: "VarCovar Rembrandt"
author: "TF"
date: "`r Sys.Date()`"
output:
  bookdown::github_document2:
    toc: True
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
library(dplyr)
library(ggplot2)

source(file = "functions/fct_report_simu_setting.R")

dfRembrandt <- readRDS("data/rembrandt/dfresvarcovar.rds")
dfBreast <- readRDS("data/breast_cancer/dfresvarcovar.rds")
```

Each dot in the following figure corresponds to one pathway.

# Rembrandt

## Proportion of significant genes

```{r}
plot_R <- fct_prop_sign_genes(dfRembrandt) + ggtitle("Rembrandt")
plot_B <- fct_prop_sign_genes(dfBreast) + ggtitle("Breast")

ggpubr::ggarrange(plot_R, plot_B)
```

Choices :

- number of genes : $NG = \{10, 50, 200\}$
- proportion of significant genes : $p_g = \{0.2, 0.5\}$

## Beta

```{r}
lsbeta_R <- fct_beta_dist(dfRembrandt, caption = "Rembrandt : Proportion of positive, negative and non significant genes")
lsbeta_B <- fct_beta_dist(dfBreast, caption = "Breast cancer : Proportion of positive, negative and non significant genes")

lsbeta_R$table_beta
lsbeta_B$table_beta
```

Proportions are roughly similar between positive and negative genes.

```{r}
plot_R <- lsbeta_R$plot_unique_beta + ggtitle("Rembrandt")
plot_B <- lsbeta_B$plot_unique_beta + ggtitle("Breast")

ggpubr::ggarrange(plot_R, plot_B)
```

```{r}
plot_R <- lsbeta_R$plot_multiple_beta + ggtitle("Rembrandt")
plot_B <- lsbeta_B$plot_multiple_beta + ggtitle("Breast")

ggpubr::ggarrange(plot_R, plot_B)
```

Choices :

- $\beta \sim \mathcal{N}(0, 0.4^2)$
- half of genes follows $\beta \sim \mathcal{N}(-0.4, 0.2^2)$ and half follows $\beta \sim \mathcal{N}(0.4, 0.2^2)$
- half of genes follows $\beta \sim \mathcal{N}(-0.8, 0.4^2)$ and half follows $\beta \sim \mathcal{N}(0.8, 0.4^2)$

## Variance

```{r}
plot_R <- fct_plot_variance(dfRembrandt) + ggtitle("Rembrandt")
plot_B <- fct_plot_variance(dfBreast, nb_filter_plot = 3) + ggtitle("Breast")

ggpubr::ggarrange(plot_R, plot_B)
```

Choice :

- Variance : $C_{jj} = 0.2$

## Correlation

```{r}
ls_R <- fct_plot_correlation(dfRembrandt, caption = "Rembrandt : Shapes of non-standard beta distribution by type of gene")
ls_B <- fct_plot_correlation(dfBreast, caption = "Breast : Shapes of non-standard beta distribution by type of gene")

plot_R <- ls_R$plot_unique + ggtitle("Rembrandt")
plot_B <- ls_B$plot_unique + ggtitle("Breast")

ggpubr::ggarrange(plot_R, plot_B)
```

```{r}
ls_R$table_shapes_correlation
ls_B$table_shapes_correlation
```

```{r}
plot_R <- ls_R$plot_multiple + ggtitle("Rembrandt")
plot_B <- ls_B$plot_multiple + ggtitle("Breast")

ggpubr::ggarrange(plot_R, plot_B)
```

Choice :

- Overall correlation follows a non-standard beta $corr \sim NSBeta(20, 20, min=-1, max =1)$
- Correlation between significant genes $corr \sim NSBeta(10, 10, min=-1, max =1)$ and between other genes $corr \sim NSBeta(25, 25, min=-1, max =1)$

## Summary Rembrandt

Taking paper notations :

- number of genes : $NG = \{10, 50, 200\}$
- proportion of significant genes : $p_g = \{0.2, 0.5\}$
- Variance : $C_{jj} = 0.2$
- Type (D) : $\beta \sim \mathcal{N}(0, 0.4^2)$
- Type (E) : half of genes follows $\beta \sim \mathcal{N}(-0.4, 0.2^2)$ and half follows $\beta \sim \mathcal{N}(0.4, 0.2^2)$
- Type (F) : half of genes follows $\beta \sim \mathcal{N}(-0.8, 0.4^2)$ and half follows $\beta \sim \mathcal{N}(0.8, 0.4^2)$
- Case (IV) : Overall correlation follows a non-standard beta $corr \sim NSBeta(20, 20, min=-1, max =1)$
- Case (V) : Correlation between significant genes $corr \sim NSBeta(10, 10, min=-1, max =1)$ and between other genes $corr \sim NSBeta(25, 25, min=-1, max =1)$

Nb of simulations to perform = $3*2*3*2 = 36$