### 2023-09-21 reporting/simu_rembrandt_breast.Rmd
update with new simu

---
title: "VarCovar Rembrandt"
author: "TF"
date: "`r Sys.Date()`"
output:
  bookdown::github_document2:
    toc: True
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
library(dplyr)
library(ggplot2)

source(file = "functions/fct_report_simu_setting.R")

dfRembrandt <- readRDS("data/rembrandt/dfresvarcovar.rds")
dfBreast <- readRDS("data/breast_cancer/dfresvarcovar.rds")
```

Each dot in the following figure corresponds to one pathway.

# Introduction

Document is separated into two parts

- First part goal is to find relevant simulation setting from Rembrandt and Breast cancer study.
- Second part will present the results from the simulation studies

# Finding the simulation setting

## Proportion of significant genes

```{r}
plot_R <- fct_prop_sign_genes(dfRembrandt) + ggtitle("Rembrandt")
plot_B <- fct_prop_sign_genes(dfBreast) + ggtitle("Breast")

ggpubr::ggarrange(plot_R, plot_B)
```

Choices :

- number of genes : $NG = \{10, 50, 200\}$
- proportion of significant genes : $p_g = \{0.2, 0.5\}$

## Beta

```{r}
lsbeta_R <- fct_beta_dist(dfRembrandt, caption = "Rembrandt : Proportion of positive, negative and non significant genes")
lsbeta_B <- fct_beta_dist(dfBreast, caption = "Breast cancer : Proportion of positive, negative and non significant genes")

lsbeta_R$table_beta
lsbeta_B$table_beta
```

Proportions are roughly similar between positive and negative genes.

```{r}
plot_R <- lsbeta_R$plot_unique_beta + ggtitle("Rembrandt")
plot_B <- lsbeta_B$plot_unique_beta + ggtitle("Breast")

ggpubr::ggarrange(plot_R, plot_B)
```

```{r}
plot_R <- lsbeta_R$plot_multiple_beta + ggtitle("Rembrandt")
plot_B <- lsbeta_B$plot_multiple_beta + ggtitle("Breast")

ggpubr::ggarrange(plot_R, plot_B)
```

Choices :

- $\beta \sim \mathcal{N}(0, 0.4^2)$
- half of genes follows $\beta \sim \mathcal{N}(-0.4, 0.2^2)$ and half follows $\beta \sim \mathcal{N}(0.4, 0.2^2)$
- half of genes follows $\beta \sim \mathcal{N}(-0.8, 0.4^2)$ and half follows $\beta \sim \mathcal{N}(0.8, 0.4^2)$

## Variance

```{r}
plot_R <- fct_plot_variance(dfRembrandt) + ggtitle("Rembrandt")
plot_B <- fct_plot_variance(dfBreast, nb_filter_plot = 3) + ggtitle("Breast")

ggpubr::ggarrange(plot_R, plot_B)
```

Choice :

- Variance : $C_{jj} = 0.2$

## Correlation

```{r}
ls_R <- fct_plot_correlation(dfRembrandt, caption = "Rembrandt : Shapes of non-standard beta distribution by type of gene")
ls_B <- fct_plot_correlation(dfBreast, caption = "Breast : Shapes of non-standard beta distribution by type of gene")

plot_R <- ls_R$plot_unique + ggtitle("Rembrandt")
plot_B <- ls_B$plot_unique + ggtitle("Breast")

ggpubr::ggarrange(plot_R, plot_B)
```

```{r}
ls_R$table_shapes_correlation
ls_B$table_shapes_correlation
```

```{r}
plot_R <- ls_R$plot_multiple + ggtitle("Rembrandt")
plot_B <- ls_B$plot_multiple + ggtitle("Breast")

ggpubr::ggarrange(plot_R, plot_B)
```

Choice :

- Overall correlation follows a non-standard beta $corr \sim NSBeta(20, 20, min=-1, max =1)$
- Correlation between significant genes $corr \sim NSBeta(10, 10, min=-1, max =1)$ and between other genes $corr \sim NSBeta(25, 25, min=-1, max =1)$

## Summary Rembrandt

Taking paper notations :

- number of genes : $NG = \{10, 50, 200\}$
- number of individuals : $N = \{50, 100\}$
- proportion of significant genes : $p_g = \{0.2, 0.5\}$
- proportion of censored genes : $c = \{0, 0.3\}$
- Variance : $C_{jj} = 0.2$

- Type (D) : $\beta \sim \mathcal{N}(0, 0.4^2)$
- Type (E) : half of genes follows $\beta \sim \mathcal{N}(-0.4, 0.2^2)$ and half follows $\beta \sim \mathcal{N}(0.4, 0.2^2)$
- Type (F) : half of genes follows $\beta \sim \mathcal{N}(-0.8, 0.4^2)$ and half follows $\beta \sim \mathcal{N}(0.8, 0.4^2)$

**Note : scenario D, E and F do not rely on correlation matrix to simulate variance covariance matrix so the those matrix are unlikely to be positive definite.**

- Type (G), (H) and (I) : same as type (D), (E), (F) where covariance matrix of $\beta$ is obtained from the correlation matrix of gene expression defined by "Case" (instead of no correlation).

- Case (IV) : Overall correlation follows a non-standard beta $corr \sim NSBeta(20, 20, min=-1, max =1)$
- Case (V) : Correlation between significant genes $corr \sim NSBeta(10, 10, min=-1, max =1)$ and between other genes $corr \sim NSBeta(25, 25, min=-1, max =1)$

**Note : Case IV and V are not guaranteed to generate positive definite matrix, after the correlation values are sampled, if the matrix is not positive definite, the nearest positive definite matrix is computed using the algorithm defined by Higham 2002 (see Matrix::nearPD).**

- Case (V) : Correlation between significant genes is $0.2$, correlation with non significant genes is $0$.

# Results of simulation study

```{r fig.height=10, fig.cap="Statistical power depending on simulation setting. N is the number of patients and NG the number of genes in the pathway."}
source("script/script_simulation/figures_simu_rembrandt_breast_v1.R")

plot_simu
```

**Note : see methods section, the covariance matrix are unlikely to be positive definite.**

```{r fig.height=10, fig.cap="Statistical power depending on simulation setting (second batch). N is the number of patients and NG the number of genes in the pathway."}
source("script/script_simulation/figures_simu_rembrandt_breast_v2.R")

plot_simu
```

Overall, all methods have close statistical power but sGBJ and Global boost test tends to be among the worst methods. There are some exceptions to note :

- Global Boost Test performs well in type F
- sGBJ performs well when N = 100 and NG = 50 which might be related to its asymptotic nature.
- In other cases, Wald and Global Test slightly outperformed other methods and Wald seems better than Global Test.

## Comprehend tests

see : https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3196970/

### Wald test

- estimate one beta by gene using cox model
- sum the wald test statistics
- H0 : variance of beta is equal to 0.

### Global test

Jelle J. Goeman and others, Testing association of a pathway with survival using gene expression data, Bioinformatics, Volume 21, Issue 9, May 2005, Pages 1950–1957, https://doi.org/10.1093/bioinformatics/bti267

- estimate one beta by gene using cox model
- makes the assumption that beta are centered on 0
- H0 : variance of beta is equal to 0.

### Global boost test

Compute the cox model negative likelihood using a boosted multivariate (all genes) cox model. P-value is computed by permutation on the negative likelihood.

### Thinking

So basically, Wald and Global are super similar, Global test would have been expected to perform better in D case because the hypothesis on beta distribution is closer. One explanation could be that this test is optimal for small deviations from null hypothesis. This might explain why is seems better than Wald where power is low and why Wald is better when power is high.

Global boost test relies on one model for all the data contrary to other models that perform one model by gene which might makes it more adapted to large number of genes. This is partially in line with good performance when N = 50.

sGBJ is similar to Wald and Global test in the sense that it is based on one cox model by feature. However, it takes into account the covariance of genes in its formula. And the data are simulated with no covariance of beta => Problem ?