### 2024-07-01 script/rembrandt/rembrant_description.Rmd
ready breast cancer scripts

---
title: "Rembrant"
author: "TF"
output: 
  bookdown::html_document2:
    toc: yes
    toc_float: true
    number_sections: FALSE
    fig_caption: TRUE
    table_caption: TRUE
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r setup, results='hide'}

library(dplyr)
library(ggplot2)
library(survival)

lapply(list.files(path = here::here("functions"), full.names = TRUE), source)

df_rembrandt <- fct_load_clean_rembrandt()

pathways = GSA::GSA.read.gmt(here::here("data/rembrandt/c6.all.v7.1.symbols.gmt"))
liste_pathways=pathways$genesets
liste_names=pathways$geneset.names
names(liste_pathways) <- liste_names
```

# Describe Rembrandt cohort

```{r desc}
##### Description
df_rembrandt |> 
  dplyr::select(AGE_RANGE,
                GENDER,
                EVENT_OS,
                OVERALL_SURVIVAL_MONTHS,
                DISEASE_TYPE) |> 
  gtsummary::tbl_summary(by = DISEASE_TYPE)
```

```{r genecount, fig.cap="Rembrandt count per patient per tumor type"}
df_rembrandt |> 
  dplyr::select(-c(AGE_RANGE,
                   GENDER,
                   EVENT_OS,
                   OVERALL_SURVIVAL_MONTHS)) |> 
  tidyr::pivot_longer(cols = -c(BIOSPECIMEN_ID, DISEASE_TYPE)) |> 
  ggplot(mapping = aes(x = BIOSPECIMEN_ID, y = value, fill = DISEASE_TYPE)) +
  geom_boxplot(outlier.shape = NA,lwd=0.1, alpha=1) +
  facet_wrap(.~DISEASE_TYPE, scales = "free_x") +
  theme_minimal() +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position = "none")+
  labs(x = "patients",
       y = "counts")

```

```{r rembrandtpca, fig.cap="First two PCA factorial plans"}
pca_model <- df_rembrandt |> 
  dplyr::select(-c(BIOSPECIMEN_ID,
                   DISEASE_TYPE,
                   AGE_RANGE,
                   GENDER,
                   EVENT_OS,
                   OVERALL_SURVIVAL_MONTHS)) |> 
  FactoMineR::PCA(ncp = 4, graph = FALSE)

pca_coord <- pca_model$ind$coord |> 
  as.data.frame() |> 
  mutate(BIOSPECIMEN_ID = df_rembrandt$BIOSPECIMEN_ID,
         DISEASE_TYPE = df_rembrandt$DISEASE_TYPE,
         .before = 1)

p1 <- plot(pca_model, choix = "ind", axes = c(1, 2))
p2 <- plot(pca_model, choix = "ind", axes = c(3, 4))

ggpubr::ggarrange(p1, p2, common.legend = TRUE)
```

Here are the kaplan meier curves for the two types of disease:

```{r kmrembrandt, fig.cap="Kaplan meier curves for Astro, Oligo"}
fit <- survfit(Surv(time=OVERALL_SURVIVAL_MONTHS, event=EVENT_OS) ~ DISEASE_TYPE,
               data = df_rembrandt)

p_type=survminer::ggsurvplot(fit,
                             data = df_rembrandt,
                             surv.median.line = "hv", # Add medians survival

                             palette = c("#8338EC", "#FFB703"),

                             # Change legends: title & labels
                             legend.title = "Type",
                             legend.labs = c("Astrocytoma", "Oligodendroglioma"),

                             conf.int = TRUE,
                             # Add risk table
                             risk.table = TRUE,
                             tables.height = 0.2,
                             tables.theme = survminer::theme_cleantable(),

                             title="Type of disease",
                             ggtheme = theme_bw() # Change ggplot2 theme
)
p_type

```

# Rembrandt pathway analysis

## Description of pathways

```{r ecdf, fig.cap="Empirical Cumulative Distribution Function of number of genes by pathway. Rembrandt study."}
lapply(liste_pathways,
       function(pathway_i) sum(pathway_i %in% colnames(df_rembrandt))) |> 
  bind_rows() |> 
  tidyr::pivot_longer(cols = everything(),
                      names_to = "pathway",
                      values_to = "nb_genes") |> 
  ggplot(mapping = aes(x = nb_genes)) +
  stat_ecdf() +
  geom_point(y = 0, shape = "|") +
  theme_minimal() +
  labs(x = "Number of genes",
       y = "Empirical Cumulative Distribution")

```

## Methods comparison

```{r importrembrandtpathway}
df_rembrandt_pathway <- lapply(list.files(here::here("results/rembrandt/"), full.names = T),
                               readRDS) |> 
  bind_rows() |> 
  select(-starts_with("time_"), -nb_genes) |> 
  tidyr::pivot_longer(cols = c("sGBJ", "GBT", "Wald", "GT"),
                      names_to = "method",
                      values_to = "p_value") |> 
  group_by(grade, method) |> 
  mutate(p_value_BH = p.adjust(p_value, method = "BH"),
         method = factor(method,
                         levels = c("sGBJ", "GBT", "GT", "Wald"),
                         labels = c("sGBJ", "Global Boost Test", "Global Test", "Wald")),
         grade = factor(grade,
                         levels = c("all", "OLIGODENDROGLIOMA", "ASTROCYTOMA"),
                         labels = c("All", "Oligodendroglioma", "Astrocytoma"))) |> 
  ungroup() |> 
  mutate(significant = as.integer(p_value < 0.05),
         significant_bh = as.integer(p_value_BH < 0.05)) |> 
  mutate(method = forcats::fct_relevel(method,
                                       "sGBJ",
                                       "Global Test",
                                       "Wald",
                                       "Global Boost Test"))

# a = df_rembrandt_pathway |> filter(grade == "Astrocytoma", method == "Wald") |> pull(p_value) |> p.adjust(method = "BH")
# b = df_rembrandt_pathway |> filter(grade == "Astrocytoma", method == "Wald") |> pull(p_value_BH)
# cor(a, b)
```

```{r nbsign, fig.height=5, fig.cap="Number of significant pathways by method."}
vec_pval_method <- c("Raw", "Benjamini-Hockberg")
names(vec_pval_method) <- c("nb_sign", "nb_sign_bh")

df_rembrandt_pathway |> 
  group_by(grade, method) |> 
  summarise(nb_sign_bh = sum(significant_bh),
            nb_sign = sum(significant)) |> 
  tidyr::pivot_longer(cols = starts_with("nb_sign")) |> 
  ggplot(mapping = aes(x = method, fill = method, y = value)) +
  geom_bar(stat = "identity") +
  facet_grid(grade ~ name, scales = "free_y", labeller = labeller(name = vec_pval_method)) +
  scale_fill_viridis_d(direction = -1) +
  theme_minimal() +
  theme(axis.text.x=element_text(angle = 45, hjust = 1)) +
  labs(x = "Method",
       y = "Nb of significant pathways",
       fill = "")
```

```{r upsetplot, fig.cap="Upset plot of the Benjamini-Hockberg p-value agreement according to the different methods", fig.height=8}
ls_plots <- df_rembrandt_pathway |> 
  filter(significant_bh == 1) |>
  group_by(pathway, grade) |> 
  summarise(methods = list(unique(as.character(method)))) |> 
  group_by(grade) |> 
  group_split() |> 
  lapply(FUN = function(df_i){
    title_i <- unique(df_i$grade)
    plot_i <- df_i |> 
      ggplot(aes(x=methods)) +
      geom_bar() +
      ggupset::scale_x_upset(n_intersections = 6, order_by = "freq") +
      theme_minimal() +
      theme(plot.margin = margin(10, 10, 10, 50)) +
      labs(title = title_i,
           x = "Methods",
           y = "Number of significant pathways")

    return(plot_i)
  })
ggpubr::ggarrange(plotlist = ls_plots, ncol = 1)
```

```{r figpvaluesmethodsrembrandt, fig.cap="Raw p-values in function of the ordered ranks of sGBJ for the 4 methods (sGBJ , global boost test, Wald test and global test), with the 5% threshold and the Benjamini Hochberg limit, computed for astrocytoma, oligodendroglioma and all patients. Nota Bene: The Benjamini Hochberg limit only applies for the sGBJ method, as the ranks are computed for sGBJ only.", fig.height=8}
nb_pathway = df_rembrandt_pathway |> 
  pull(pathway) |> 
  unique() |> 
  length()

df_sGBJ_rank <- df_rembrandt_pathway |> 
  group_by(grade) |> 
  filter(method == "sGBJ") |> 
  mutate(sGBJ_rank = rank(p_value)) |> 
  ungroup() |> 
  select(pathway, grade, sGBJ_rank)

df_BH_limit <- data.frame(sGBJ_rank = seq(1:nb_pathway)) |> 
  mutate(p_value = 0.05/nb_pathway * sGBJ_rank)

df_rembrandt_pathway |> 
  left_join(df_sGBJ_rank, by = c("pathway", "grade")) |> 
  ggplot(mapping = aes(x = sGBJ_rank, y = p_value)) +
  geom_point(mapping = aes(color = method)) +
  scale_color_viridis_d(direction = -1) +
  ggnewscale::new_scale_color() +
  geom_hline(mapping = aes(color = "5% threshold",
                           linetype = "5% threshold", yintercept = 0.05),
                         linewidth = 1) +
  geom_line(data = df_BH_limit,
            linewidth = 1,
            mapping = aes(color = "BH limit", linetype = "BH limit")) +
  scale_color_manual(values = c("red", "black")) +
  scale_linetype_manual(values = c(2,1)) +
  labs(color = "", linetype = "") +

  facet_grid(grade ~ ., scales = "free_y") +
  scale_y_log10() +
  annotation_logticks(sides = "l") +
  theme_minimal() +
  labs(x = "SGBJ rank", y = "Raw p-value")
```

NB about GT test : https://www.bioconductor.org/packages/release/bioc/vignettes/globaltest/inst/doc/GlobalTest.pdf : "Because permutations require an exchangeable null hypothesis, such a permutation p-value is only available for the linear model and for the exchangeable null hypotheses ~1 and ~0 in other models."